WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-07-31 ms.tex
Add MBTA reference

The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries. This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}; \aclu{MBTA}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities. Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.

>>>>

The final science run of the initial LIGO and Virgo instruments saw the first joint search for \ac{GW} and electromagnetic emission from compact binaries. This involved several advances in the \ac{GW} data analysis \citep{CBCLowLatency}, including the first real\nobreakdashes-time matched\nobreakdashes-filter detection pipeline (\acs{MBTA}, \aclu{MBTA}; \cite{LIGOVirgoInspiralPipelineComparison}), a semi\nobreakdashes-coherent, ad hoc rapid triangulation code (Timing++), and the first version of a rigorous Bayesian \ac{MCMC} parameter estimation code (LALINFERENCE; \citealt{S6PE})---all in service of the first search for X\nobreakdashes-ray \citep{SwiftFollowup} and optical \citep{OpticalImageAnalysis} counterparts of \ac{GW} triggers, by a consortium of facilities. Despite the technical achievements in the \ac{GW} data analysis, there was an undesirable tradeoff between the speed as well as accuracy of the rapid localization and the full parameter estimation: the former could analyze a detection candidate in minutes, whereas the latter took half a day; the latter decreased the area on the sky by a factor of 1/20 over the former but took 1000 times as long to run \cite{SiderySkyLocalizationComparison}.
